### 2019-08-19 code/analysis2.Rmd

---
title: "Analysis - Part 2"
subtitle: "PsyFaKo Open Science Survey"
author: "Johannes Brachem"
date: "Last update: `r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    number_sections: yes
    theme: journal
    toc: yes
    toc_depth: 2
    toc_float: 
      collapsed: false
---

```{r include=FALSE}
# knitr::opts_chunk$set(echo=FALSE)
```

# Header
## Load Packages
```{r packages}
library(tidyverse)
```

## Session Info
```{r session info}
sessionInfo()
```

## Import Data

```{r import data}
main <- read_csv("data/main.csv") %>%
  filter(field_of_study_mc == "Psychology" | study_degree_field == "Psychology")
stats <- readRDS("data/stats_software.rds")
```

```{r}
# global variables
n_participants <- nrow(main)
```

# Question Block 3: Teaching Coverage, Attitudes

## Specific QRPs in teaching
Which of these research practices have been talked about in a critical way in the course of your studies?
```{r QRPs mentioned critically}
q_qrp_t.spec <- main %>% 
  mutate(id = 1:nrow(.)) %>% 
  select(id, study_stage_mc, semester, specific_qrps_teaching) %>% 
  separate(specific_qrps_teaching, into = paste0("qrp.t", 1:11)) %>% 
  gather(qrp.t1:qrp.t11, key = "answer_number", value = "qrp_id") %>% 
  drop_na(qrp_id) %>% 
  mutate(qrp = dplyr::recode(qrp_id, 
                             "1" = "SR of Variables", 
                             "2" =    "No Sample Planning", 
                             "3" =     "Flexible Sample Size", 
                             "4" =    "SR of Conditions", 
                             "5" =   "Flexible Analysis", 
                             "6" =    "Flexible Exclusion", 
                             "7" =    "SR of Hypotheses", 
                             "8" =   "HARKing", 
                             "9" =    "Rounding p-Values", 
                             "10" =   "None", 
                             "11" =   "Not Sure")) %>% 
  select(-answer_number, -qrp_id)

q_qrp_t.spec %>% 
  group_by(qrp) %>% summarise(n = n()) %>% ungroup() %>% 
  mutate(share = n / n_participants * 100) %>% arrange(share %>% desc())
```

## Replication Crisis in any lecture

Did you hear about the replication crisis in a lecture?
```{r RC in any lecture}
rc_teaching_general <- main %>% 
  # summarise semesters into year-chunks, 
  # filter(study_stage_mc != "Other") %>% 
  mutate(semester = ifelse(semester %% 2 == 0, semester - 1, semester)) %>% 
  mutate(semester = ifelse(semester > 10, 11, semester)) %>% 
  mutate(semester = factor(semester, 
                           labels = c("1/2", "3/4", "5/6", 
                                      "7/8", "9/10", ">10"))) %>%

  # select relevant variables and summarise
  select(rc_teaching, semester, study_stage_mc) %>% 
  group_by(rc_teaching, semester, study_stage_mc) %>% 
  summarise(n = n()) %>% 
  ungroup() %>%

  # number of participants per semester and study stage
  group_by(semester, study_stage_mc) %>% 
  mutate(n_semester = sum(n)) %>% 
  ungroup() %>%

  # share with respect to students in the same study stage and semester
  mutate(share = (n / n_semester) * 100)

# print
rc_teaching_general %>% 
  arrange(study_stage_mc) %>%  
  print(n = nrow(.))
```

## Replication Crisis in which lectures
In which lectures did you hear about the replication crisis?
```{r RC in which lectures}
q_rc_t.spec <- main %>% 
  mutate(id = 1:nrow(.)) %>% 
  separate(rc_teaching_specific, into = paste0("t", 1:11)) %>% 
  select(id, study_stage_mc, semester, t1:t11) %>% 
  gather(t1:t11, key = "answer_number", value = "lecture_id") %>% 
  drop_na(lecture_id) %>% 
  mutate(lecture = dplyr::recode(lecture_id, 
                                 "1" = "General Psychology",
                                 "2" = "Work, Organizational and Economic Psychology",
                                 "3" = "Biological- / Neuropsychology",
                                 "4" = "Diagnostics",
                                 "5" = "Differential Psychology",
                                 "6" = "Introductory Class",
                                 "7" = "Developmental Psychology",
                                 "8" = "Health Psychology",
                                 "9" = "Clinical Psychology",
                                 "10" = "Media- / Communication Psychology",
                                 "11" = "Methods, Stats, Expra",
                                 "12" = "Pedagogical Psychology",
                                 "13" = "Forensic",
                                 "14" = "Social Psychology",
                                 "15" = "Other")) %>% 
  select(-answer_number, -lecture_id)

q_rc_t.spec %>% 
  group_by(lecture) %>% summarise(n = n()) %>% ungroup() %>% 
  mutate(share = n / n_participants * 100) %>% arrange(share %>% desc())
```

## Attitudes towards replication crisis and open science

```{r}
# table of answers for the three questions
# 1: "not at all"
# 5: "very much"
main %>% 
  select(interest, importance, felt_information) %>% 
  map(table)
```

# Question Block 4: Other

## Statistical Software

```{r}
# print
stats %>% print(n = nrow(.))
```

## Open Science Initiatives
Do you know of an open science initiative at your university?
```{r}
# light wrangling
nosi_aggregation <- main %>% select(nosi, uni_current, nosi_at_uni) %>% 
  group_by(uni_current, nosi_at_uni, nosi) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  mutate(nosi = factor(nosi, levels = c("No", "Not Sure", "Yes"))) %>% 
  arrange(desc(nosi_at_uni))

# compute shares
nosi_agg2 <- nosi_aggregation %>% 
  group_by(uni_current) %>% 
  mutate(n_answers = sum(n)) %>% 
  filter(nosi == "Yes") %>% 
  mutate(share_yes = (n / n_answers) * 100) %>% 
  select(uni_current, nosi_at_uni, n_answers, share_yes) %>% 
  arrange(desc(share_yes))

# print
nosi_agg2
```

Have you heard of open science events (e.g. workshops or talks) at your university=
```{r}
nosi_events <- main %>% select(nosi_events_mc) %>% 
  group_by(nosi_events_mc) %>% 
  summarise(n = n())

nosi_events
```

## Block 3
```{r}
block3 <- list(qrps_teaching = q_qrp_t.spec,
               rc_teaching_general = rc_teaching_general,
               rc_teaching_specific = q_rc_t.spec,
               nosi_data = nosi_agg2,
               nosi_events = nosi_events
               )

saveRDS(block3, "data/block3_results.rds")
```