### 2019-08-19 manuscript/manuscript-jo.tex

\documentclass[man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={The title},
            pdfauthor={First Author~\& Ernst-August Doelle},
            pdfkeywords={keywords},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

  \title{The title}
    \author{First Author\textsuperscript{1}~\& Ernst-August
Doelle\textsuperscript{1,2}}
    \date{}

\shorttitle{Title}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Wilhelm-Wundt-University\\\textsuperscript{2} Konstanz Business School}
\keywords{keywords\newline\indent Word count: X}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

\usepackage{longtable}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}

\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

\usepackage{lineno}

\linenumbers
\usepackage[ngerman]{babel}

\authornote{Add complete departmental affiliations for each
author here. Each new line herein must be indented, like this line.

Enter author note here.

Correspondence concerning this article should be addressed to First
Author, Postal address. E-mail:
\href{mailto:my@email.com}{\nolinkurl{my@email.com}}}

\abstract{
One or two sentences providing a \textbf{basic introduction} to the
field, comprehensible to a scientist in any discipline.

Two to three sentences of \textbf{more detailed background},
comprehensible to scientists in related disciplines.

One sentence clearly stating the \textbf{general problem} being
addressed by this particular study.

One sentence summarizing the main result (with the words ``\textbf{here
we show}'' or their equivalent).

Two or three sentences explaining what the \textbf{main result} reveals
in direct comparison to what was thought to be the case previously, or
how the main result adds to previous knowledge.

One or two sentences to put the results into a more \textbf{general
context}.

Two or three sentences to provide a \textbf{broader perspective},
readily comprehensible to a scientist in any discipline.

}

\begin{document}
\maketitle

\section{Methoden}\label{methoden}

We report how we determined our sample size, all data exclusions (if
any), all manipulations, and all measures in the study.

\subsection{Teilnehmende}\label{teilnehmende}

\subsection{Prozedur}\label{prozedur}

\subsection{Datenanalyse}\label{datenanalyse}

We used for all our analyses.

\section{Ergebnisse}\label{ergebnisse}

Aus den Angaben der Teilnehmenden ergab sich ein mittlerer Einsatz von
1.32 fragwürdigen Forschungspraktiken (SD = 1.49) pro Projekt. Wie in
\autoref{fig:plot02} zu sehen ist, zeigten sich dabei Unterschiede
zwischen den abgefragten Projekten. Besonders viele fragwürdige
Praktiken scheinen demnach mit durchschnittlich 1.41 Praktiken pro
Projekt im Expra verwendet zu werden, während Masterarbeiten mit
durchschnittlich 0.54 verwendeten fragwürdigen Praktiken den niedrigsten
Wert aufweisen.

\begin{figure}
\centering
\includegraphics{manuscript-jo_files/figure-latex/plot02-1.pdf}
\caption{\label{fig:plot02} extbf\{Anzahl eingesetzter fragwürdiger
Forschungspraktiken pro Projekt.\} Schwarze Punkte stellen den
Mittelwert dar. Die Anzahl ist ganzzahlig, die Punkte wurden mit
zufälliger Streuung versehen.}
\end{figure}

In \autoref{fig:plot0305} ist die Einsatzhäufigkeit der einzelnen
abgefragten Forschungspraktiken dargestellt. Als am häufigsten
eingesetzte fragwürdige Forschungsmethoden zeigten sich fehlende
Stichprobenplanung und selektives Berichten abhängiger Variablen (34.93,
bzw. 23.49 \% der Projekte betroffen). Die von unseren Teilnehmenden am
seltensten eingesetzten Methoden waren das Abrunden von p-Werten und das
Nacherheben von Versuchspersonen zum Zweck der Herbeiführung
statistischer Signifikanz (2.08, bzw. 6.57 \% der Projekte betroffen).

\begin{figure}
\centering
\includegraphics{manuscript-jo_files/figure-latex/plot0305-1.pdf}
\caption{\label{fig:plot0305}Caption}
\end{figure}

\section{Diskussion}\label{diskussion}

\newpage

\section{References}\label{references}

\begingroup
\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}

\hypertarget{refs}{}

\endgroup

\end{document}