### 2019-09-08 code/analysis.Rmd
big update

---
title: "Descriptive Analysis"
subtitle: "PsyFaKo Open Science Survey"
author: "Johannes Brachem"
date: "Last update: `r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    number_sections: yes
    theme: journal
    toc: yes
    toc_depth: 2
    toc_float: 
      collapsed: false
---

# Header
## Load Packages

```{r}
source("code/helpers/helper_functions.R")
```

## Import Data

```{r import data}
wide <- read_csv("data/wide.csv") %>% 
  only_psych() %>% filter(emp_experience == "Yes")

long <- read_csv("data/long.csv") %>% 
  only_psych() %>% filter(emp_experience == "Yes")

wide_all <- read_csv("data/wide.csv") %>% only_psych()
stats <- readRDS("data/stats_software.rds")
```

# Question Block 2: Research Practices

```{r}
n_projects_total <- long %>% 
  distinct(id, n_projects) %>% 
  summarise(n = sum(n_projects, na.rm = TRUE)) %>% 
  pull(n)
```

## Preregistered Analyses

**Note**: In our preregistration, we only specified analyses for questionable research practices. Because it might be of interest, we also calculate the same figures for the two open research practices that we asked about. One should be very careful to compare both types of practices here directly, because the number of questionable research practices that we asked about was a lot higher (9), than the number of open practices that we asked about (2). So, e.g. a difference of 1.3 questionable practices applied per project on average, compared to only 0.5 open practices is almost certainly confounded by the number of specific practices that were assessed.

### Full Aggregation

#### Mean Number of QRPs and ORPs applied per Project per Participant
The chunk below first sums up, how many QRPs and ORPs were totally applied by each participant. Then, theses sums are each divided by the total number of projects conducted by each participant, returning the average number of QRPs / ORPs applied per project.

```{r eval=FALSE, include=FALSE}
# Total scores: Total number of RP applications divided by total number of projects 
# (not preregistered, not included in output)

q01.0 <- long %>% 
  group_by(type_of_practice) %>% 
  summarise(n = sum(rp_applied, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(n_projects = n_projects) %>% 
  mutate(practices_per_project = n / n_projects)

q01.0
```

```{r}
# Mean over individual scores: Number of RP applications per project,
# averaged across individuals (preregistered)

q01 <- long %>% 
  # number of applied practices (per individual and type of project)
  group_by(id, type_of_practice, n_projects) %>% 
  summarise(applied_practices = sum(rp_applied, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(practices_per_project = applied_practices / n_projects) %>%

  # average over individuals
  group_by(type_of_practice) %>% 
  summarise(m = mean(practices_per_project),
            sd = sd(practices_per_project),
            med = median(practices_per_project),
            minimum = min(practices_per_project),
            maximum = max(practices_per_project))

q01
```

### Project Aggregation

How many open/questionable research practices were applied on average per project?

```{r}
q02 <- long %>% 
  group_by(id, type_of_practice, project, project_conducted) %>% 
  summarise(n_practice_appl = sum(rp_applied)) %>% 
  ungroup() %>% 
  group_by(type_of_practice, project) %>% 
  summarise(mean = mean(n_practice_appl, na.rm = TRUE), # mean number of applied practices per project
            sd = sd(n_practice_appl, na.rm = TRUE), # sd for number of applied practices per project
            n_conducted = sum(project_conducted), # number of conducted projects
            n_applied = sum(n_practice_appl, na.rm=TRUE), # number of practice applications
            med = median(n_practice_appl, na.rm = T),
            minimum = min(n_practice_appl, na.rm = T),
            maximum = max(n_practice_appl, na.rm = T))

q02
```

### Practice Aggregation
For every practice, we compute the share of participants who applied that practice *at least one* in any project. We only include participants who have conducted at least one empirical research project.

```{r}
q03 <- long %>% 
  filter(rp_applied) %>% 
  group_by(practice, type_of_practice) %>% 
  summarise(n = n_distinct(id)) %>% 
  mutate(n_vp = nrow(wide)) %>% 
  mutate(share = (n / n_vp)*100) %>% 
  arrange(share %>% desc()) %>% 
  identity()

# print
q03
```

### Detailed Practice Aggregation

```{r}
q04.n_projects <- long %>% 
  ungroup() %>% 
  distinct(id, project, project_conducted) %>% 
  group_by(project) %>% 
  summarise(n = sum(project_conducted)) %>% 
  mutate(order = c(1, 5, 2, 3, 4)) %>% 
  arrange(order) %>% 
  arrange(project)

q04.d <- long %>% arrange(project) %>% 
  mutate(q04.n = rep(q04.n_projects$n, each = nrow(long) / 5))

q04 <- q04.d %>% 
  ungroup() %>% 
  group_by(practice, type_of_practice, project, q04.n) %>% 
  # group_by(practice) %>% 
  summarise(n = sum(rp_applied, na.rm = TRUE)) %>% ungroup() %>% 
  # group_by(practice, type_of_practice, project, q04.n) %>% 
  mutate(share = (n / q04.n)*100) %>% 
  arrange(type_of_practice) %>% 
  identity()

# print
q04 %>% print(n = nrow(.))
```

## Platforms for Preregistration
`prereg_where` contains the answers to the question "Which platform did you use for your preregistration?". Multiple answers where possible. They are coded:

answer | meaning
-------|----------
1 | preregistration given to lecturer
2 | aspredicted.org
3 | osf.io
4 | journal
5 | other
6 | I never preregistered

```{r}
prereg_where <- wide %>% 
  pull(prereg_where)

prereg_counts <- as.character(1:6) %>% 
  map_dbl(function(x) prereg_where %>% str_count(x) %>% sum(na.rm=T))

names_prereg_where <- c("lecturer", "aspredicted.org", "osf.io", 
                        "journal", "other", "never preregistered")

prereg_where_data <- tibble(names_prereg_where, prereg_counts) %>% 
  mutate(share = prereg_counts / nrow(wide))
```

```{r}
prereg_where_data %>% 
  ggplot(aes(x = fct_reorder(names_prereg_where, 
                             prereg_counts), 
             y = prereg_counts)) +
  geom_bar(stat = "identity", aes(fill = names_prereg_where)) +
  labs(x = "Platform", y = "Number of registrations",
       title = "Platforms used for preregistration",
       fill = "Platform") +
  coord_flip() +
  theme_classic() +
  theme(legend.position = "none") +
  # theme_fivethirtyeight() +
  # ggthemes::theme_solarized() + 
  NULL
```

## Non-registered Analyses

### Share of practice applications
For each practice, we look at the share of projects in which the respective practice was applied.
```{r}
n_proj <- long %>% distinct(id, project, project_conducted) %>% 
  summarise(n = sum(project_conducted)) %>% pull(n)

q05 <- long %>% 
  # filter(project %contains% "thesis.bsc" | project %contains% "thesis.msc") %>% 
  filter(rp_applied) %>% 
  group_by(practice, type_of_practice) %>% 
  summarise(n_applied = n()) %>% 
  mutate(n_projects = n_proj) %>% 
  mutate(share = (n_applied / n_projects)*100) %>% 
  arrange(share %>% desc())

q05
```

### Open Science Initiatives

```{r}
nosi_q <- long %>% 
  # number of applied practices (per individual and type of project)
  group_by(id, type_of_practice, n_projects, nosi_at_uni) %>% 
  summarise(applied_practices = sum(rp_applied, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(practices_per_project = applied_practices / n_projects)

  # average over individuals
nosi_q %>%
  group_by(type_of_practice, nosi_at_uni) %>% 
  summarise(m = mean(practices_per_project),
            sd = sd(practices_per_project))
```

# Save Results
```{r}
block2 <- list(q01 = q01, 
                    q02 = q02, 
                    q03 = q03, 
                    q04 = q04, 
                    q05 = q05,
               prereg_data = prereg_where_data)

saveRDS(block2, "data/block2_results.rds")
```

# Question Block 3: Teaching Coverage, Attitudes
```{r}
# global
n_participants <- nrow(wide_all)
```

## Specific QRPs in teaching
Which of these research practices have been talked about in a critical way in the course of your studies?
```{r QRPs mentioned critically}
q_qrp_t.spec <- wide_all %>% 
  mutate(id = 1:nrow(.)) %>% 
  select(id, study_stage_mc, semester, specific_qrps_teaching) %>% 
  separate(specific_qrps_teaching, into = paste0("qrp.t", 1:11)) %>% 
  gather(qrp.t1:qrp.t11, key = "answer_number", value = "qrp_id") %>% 
  drop_na(qrp_id) %>% 
  mutate(qrp = dplyr::recode(qrp_id, 
                             "1" = "SR of Variables", 
                             "2" =    "No Sample Planning", 
                             "3" =     "Flexible Sample Size", 
                             "4" =    "SR of Conditions", 
                             "5" =   "Flexible Analysis", 
                             "6" =    "Flexible Exclusion", 
                             "7" =    "SR of Hypotheses", 
                             "8" =   "HARKing", 
                             "9" =    "Rounding p-Values", 
                             "10" =   "None", 
                             "11" =   "Not Sure")) %>% 
  select(-answer_number, -qrp_id)

q_qrp_t.spec_share <- q_qrp_t.spec %>% 
  group_by(qrp) %>% summarise(n = n()) %>% ungroup() %>% 
  mutate(share = n / n_participants * 100) %>% arrange(share %>% desc())
```

## Replication Crisis in any lecture

Did you hear about the replication crisis in a lecture?
```{r RC in any lecture}
rc_teaching_general <- wide_all %>% 
  # summarise semesters into year-chunks, 
  # filter(study_stage_mc != "Other") %>% 
  mutate(semester = ifelse(semester %% 2 == 0, semester - 1, semester)) %>% 
  mutate(semester = ifelse(semester > 10, 11, semester)) %>% 
  mutate(semester = ifelse(is.na(semester), "not applicable", semester)) %>% 
  mutate(semester = factor(semester, levels = c(1, 3, 5, 7, 9, 11, "not applicable"),
                           labels = c("1/2", "3/4", "5/6", 
                                      "7/8", "9/10", ">10", "not applicable"))) %>%

  # select relevant variables and summarise
  select(rc_teaching, semester, study_stage_mc) %>% 
  group_by(rc_teaching, semester, study_stage_mc) %>% 
  summarise(n = n()) %>% 
  ungroup() %>%

  # number of participants per semester and study stage
  group_by(semester, study_stage_mc) %>% 
  mutate(n_semester = sum(n)) %>% 
  ungroup() %>%

  # share with respect to students in the same study stage and semester
  mutate(share = (n / n_semester) * 100)

# print
rc_teaching_general %>% 
  arrange(study_stage_mc) %>%  
  print(n = nrow(.))
```

Replication crisis in lectures by university
```{r}
rc_teaching_unis <- wide_all %>%
  group_by(uni_current) %>% mutate(n_uni = n()) %>% ungroup() %>% 
  group_by(uni_current, rc_teaching) %>% 
  mutate(n = n()) %>% ungroup() %>% 
  mutate(share = n / n_uni) %>% 
  distinct(uni_current, n_uni, rc_teaching, n, share) %>% arrange(uni_current)

rc_teaching_unis %>% 
  mutate(share = (share*100) %>% round(1),
         share = paste0("(", share, " %)")) %>% 
  unite(n, share, col = "n", sep = " ") %>% 
  spread(key = rc_teaching, value = n, fill = "0 (0 %)")

rc_teaching_unis.out <- rc_teaching_unis %>% 
  mutate(rc_teaching = factor(rc_teaching, levels = lab.rc_lectures,
                              labels = c("yes", "no", "not_sure"))) %>% 
  unite(n, share, col = "n_share", sep = "_") %>% 
  spread(key = rc_teaching, value = n_share, fill = "0_0") %>% 
  separate(yes, into = c("yes.n", "yes.share"), sep = "_") %>% 
  separate(no, into = c("no.n", "no.share"), sep = "_") %>% 
  separate(not_sure, into = c("not_sure.n", "not_sure.share"), sep = "_") %>% 
  arrange(desc(yes.share)) %>% 
  filter(n_uni >= 30) %>% 
  mutate_at(vars(matches("share")), as.numeric) %>% 
  mutate_at(vars(matches("share")), cop2)
```

## Replication Crisis in which lectures
In which lectures did you hear about the replication crisis?
```{r RC in which lectures}
q_rc_t.spec <- wide_all %>% 
  mutate(id = 1:nrow(.)) %>% 
  separate(rc_teaching_specific, into = paste0("t", 1:11)) %>% 
  select(id, study_stage_mc, semester, t1:t11) %>% 
  gather(t1:t11, key = "answer_number", value = "lecture_id") %>% 
  drop_na(lecture_id) %>% 
  mutate(lecture = dplyr::recode(lecture_id, 
                                 "1" = "General Psychology",
                                 "2" = "Work, Organizational and Economic Psychology",
                                 "3" = "Biological- / Neuropsychology",
                                 "4" = "Diagnostics",
                                 "5" = "Differential Psychology",
                                 "6" = "Introductory Class",
                                 "7" = "Developmental Psychology",
                                 "8" = "Health Psychology",
                                 "9" = "Clinical Psychology",
                                 "10" = "Media- / Communication Psychology",
                                 "11" = "Methods, Stats, Expra",
                                 "12" = "Pedagogical Psychology",
                                 "13" = "Forensic",
                                 "14" = "Social Psychology",
                                 "15" = "Other")) %>% 
  select(-answer_number, -lecture_id)

q_rc_t.spec.share <- q_rc_t.spec %>% 
  group_by(lecture) %>% summarise(n = n()) %>% ungroup() %>% 
  mutate(share = n / n_participants * 100) %>% arrange(share %>% desc())
```

## Attitudes towards replication crisis and open science

```{r}
# table of answers for the three questions
# 1: "not at all"
# 5: "very much"
wide_all %>% 
  select(interest, importance, felt_information) %>% 
  map(table)
```

# Question Block 4: Other

## Statistical Software

```{r}
# print
stats %>% print(n = nrow(.))
```

## Open Science Initiatives
Do you know of an open science initiative at your university?
```{r}
# light wrangling
nosi_aggregation <- wide_all %>% select(nosi, uni_current, nosi_at_uni) %>% 
  group_by(uni_current, nosi_at_uni, nosi) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  mutate(nosi = factor(nosi, levels = c("No", "Not Sure", "Yes"))) %>% 
  arrange(desc(nosi_at_uni))

# compute shares
nosi_agg2 <- nosi_aggregation %>% 
  group_by(uni_current) %>% 
  mutate(n_answers = sum(n)) %>% 
  filter(nosi == "Yes") %>% 
  mutate(share_yes = (n / n_answers) * 100) %>% 
  select(uni_current, nosi_at_uni, n_answers, share_yes) %>% 
  arrange(desc(share_yes))

# print
nosi_agg2
```

Have you heard of open science events (e.g. workshops or talks) at your university=
```{r}
nosi_events <- wide_all %>% select(nosi_events_mc) %>% 
  group_by(nosi_events_mc) %>% 
  summarise(n = n())

nosi_events
```

## Block 3
```{r}
block3 <- list(qrps_teaching = q_qrp_t.spec,
               qrps_teaching_share = q_qrp_t.spec_share,
               rc_teaching_general = rc_teaching_general,
               rc_teaching_unis = rc_teaching_unis.out,
               rc_teaching_specific = q_rc_t.spec,
               rc_teaching_specific_share = q_rc_t.spec.share,
               nosi_data = nosi_agg2,
               nosi_events = nosi_events
               )

saveRDS(block3, "data/block3_results.rds")
```