### 2019-09-08 code/analysis2.Rmd

---
title: "Analysis - Part 2"
subtitle: "PsyFaKo Open Science Survey"
author: "Johannes Brachem"
date: "Last update: `r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    number_sections: yes
    theme: journal
    toc: yes
    toc_depth: 2
    toc_float: 
      collapsed: false
---

```{r include=FALSE}
# knitr::opts_chunk$set(echo=FALSE)
```