### 2019-09-08 code/flextable_test.Rmd

---
title: "Untitled"
author: "Johannes Brachem"
date: "1 9 2019"
output:
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(flextable)
library(tidyverse)
source("code/helper_functions.R")
```

```{r}
d <- mtcars %>% mutate(mpg = co(mpg)) %>% 
  mutate_at(c(3,4), cop)

myft <- flextable(
  head(d), 
  col_keys = c("am", "carb", "gear", "mpg", "drat" ))
myft %>% align(j = 3:4, align = "right", part = "all") %>% 
  colformat_num(col_keys =c("am", "carb", "gear", "drat"), digits = 2) %>% 
  theme_booktabs() %>% 
  autofit() %>%
  fix_border_issues() %>% 
  footnote(i = 1, j = 1, 
           value = as_paragraph("This is a footnote"),
           ref_symbols = "a")

```